---
title: "Covid91 vaccine study Final2024"
author: "Mariahlynn"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: " summer 2024" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$: The Covid921 vaccine is not effective in males

$H_A$: The Covid921 vaccine is effective in males 

### Methods

Treatment is a categorical variable  
MALES is also a categorical variable  
We use the cat~cat helper file  

### Descriptive Results 

### Graphical Results 
```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="Percent")
```

This bar charts shows that MALES who received the vaccine were less likely to get Covid921 in comparison to those who received a placebo. Therefore about 50% of MALES who received either the vaccine or placebo did not contract the virus. 


### Numerical Results 
```{r}
table1 <- xtabs(~treatment + infected, data=males)
rowPerc(table1)
colPerc(table1)
```

This shows that 74% of the males who took the placebo got covid19, as were 26% males who recieved the vaccine got the virus. 

### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi-squared test and Fisher's Exact Test both show a P-value of 0.00000000000000022, which is less than 0.05. 
Therefore, we reject the null hypothesis. A statistically significant effect of the vaccine.  We see that MALES are 2.8 times more likely to contract Covid921 with the placebo compared to the vaccine.

## Females

$H_0$: The Covid921 vaccine is not effective in females

$H_A$: The Covid921 vaccine is effective in females

### Methods


Female is CAT
Treatment is CAT
Helper File we use
CAT ~ CAT

### Descriptive Results

#### Graphical Results


```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females,type="percent")

```

This graph indicates that females who received the placebo had a higher infection rate compared to those who were vaccinated. However, around 50% of females who got the placebo remained uninfected, while about 51% of those who received the vaccine also stayed uninfected.

#### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected,data=females)
rowPerc(table1)
colPerc(table1)
```

The numerical results indicate that the COVID-19 vaccine is effective in reducing the likelihood of contracting COVID-19 compared to the placebo. The vaccine decreases the odds of getting COVID-19 by about 46.9% compared to the placebo. Additionally, the relative risk reduction is around 45.5%, meaning the vaccine lowers the risk of infection by approximately 45.5% compared to the placebo.

#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-square test, with a p-value of 0.00000001914, and Fisher’s Exact Test, with a p-value of 0.00000001246, both show that the difference in infection rates between the vaccine and placebo groups is statistically significant. The odds ratio of 1.8562 means that the likelihood of getting COVID-19 with the placebo is about 1.86 times higher than with the vaccine. The 95% confidence interval, which ranges from 1.49 to 2.32, further supports the vaccine's effectiveness by suggesting that the true odds ratio falls within this range.


## LGBTQ

$H_0$: The Covid921 vaccine is not effective in LGBTQ Community.

$H_A$: The Covid921 vaccine is effective in LGBTQ Community.


### Methods 

Treatment is a categorical variable. LGBTQ is a categorical variable,so we use is cat~cat helper file   

### Descriptive Results 

### Graphical Results

```{r}
LGBTQ = filter(FinalData, LGBTQ=="gay")
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")
```

The bar charts illustrate that LGBTQ individuals who were vaccinated had a lower likelihood of contracting Covid-19 compared to those who received a placebo. Nevertheless, approximately 45% of the LGBTQ individuals who received the placebo remained uninfected, while around 50% of those who received the vaccine were uninfected.   



### Numerical Descriptive Results

```{r}
table1 <- xtabs(~treatment + infected, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

The numerical data reveal that roughly 80% of LGBTQ participants who received the placebo contracted Covid-19, compared to only 20% of those who received the vaccine. This indicates that the vaccine lowers the relative risk of infection by about 55.2%.


### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test yields a p-value of 0.000002266, while Fisher's Exact Test gives a p-value of 0.0000007269, both of which are well below the 0.05 area. Therefore, we reject the null hypothesis. This result indicates a statistically significant effect of the vaccine, demonstrating that the odds of contracting Covid921 are substantially lower with the vaccine compared to the placebo.


## DrugUsers 

$H_0$: The Covid921 vaccine is not effective in Drug users 

$H_A$: The Covid921 vaccine is effective in Drug users

### Methods 

Treatment is a categorical variable   
DRUGGIES is a categorical variable   
So we use is cat~cat helper file    

### Descriptive Results 

### Graphical Results

```{r}
DrugUser=filter(FinalData,FinalData$DrugUser=="yes")

barchartGC(~treatment + infected,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser,type="percent")
```

The bar charts reveal an unexpected finding. The drug users who received the placebo were less likely to contract Covid921 than those who received the vaccine. However, it's important to note that around 50% of drug users in both groups did not contract the virus.


### Numerical Descriptive results  Results

```{r}
table1 <- xtabs(~treatment + infected, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

The numerical data show a 3.28% difference in infection rates: 3.61% of those who received the vaccine contracted Covid921, compared to just 0.33% of those who received the placebo. This indicates that the vaccine may not be effective in drug users. Typically, a vaccine should reduce the risk of contracting Covid921, rather than increase it.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test yields a P-value of 0.0000000001414, while Fisher's Exact Test provides a P-value of 0.000000000003261. Both p-values are well below the 0.05 threshold, suggesting a significant difference between the vaccine and placebo groups. Statistically, we would reject the null hypothesis. Therefore, the data indicates an unexpected increase in infection risk among drug users who received the vaccine. Consequently, despite the statistical significance, we fail to reject the null hypothesis based on these findings.

# Individual Conclusions

## MALES

In the study, the data showed that males who got the vaccine were much less likely to get Covid921, with only 26% of them infected compared to 74% of those who got the placebo. The analysis using chi-squared and Fisher's Exact Test produced very low p-values (0.00000000000000022), far below 0.05, which means we reject the null hypothesis. Therefore, we conclude that the vaccine is effective in reducing the risk of Covid921 for males.

Chi-Square Test p-value: 2.2

Fisher’s Exact Test p-value: 2.2

Odds Ratio: 2.82 (95% CI: [2.21, 3.61])


## FEMALES

In the study involving females, the vaccine significantly reduced the incidence of Covid921 compared to the placebo. The vaccine cut the risk of infection by around 46.9%. Both chi-squared and Fisher's Exact Test showed p-values far below 0.05 (0.00000001914 and 0.00000001246, respectively), leading us to reject the null hypothesis. This strongly suggests that the Covid921 vaccine is effective in decreasing infections among females.

Chi-Square Test p-value: 1.914

 
Fisher’s Exact Test p-value: 1.246

 
Odds Ratio: 1.86 (95% CI: [1.49, 2.32])


## LGBTQ

In the study involving the LGBTQ community, both the graphical and numerical data showed a significant decrease in Covid921 cases among those who received the vaccine. The vaccine reduced the risk by about 55.2% compared to the placebo. The very low p-values from the chi-squared test (0.000002266) and Fisher's Exact Test (0.0000007269) support rejecting the null hypothesis, confirming that the vaccine is effective in lowering the risk of contracting Covid921 within the LGBTQ community.

Chi-Square Test p-value: 2.266

 
Fisher’s Exact Test p-value: 7.269

Odds Ratio: 0.24 (95% CI: [0.17, 0.33])



## DRUG USERS

In the case of drug users, the data showed a higher infection rate among those who received the vaccine (3.61%) compared to those who got the placebo (0.33%), which was unexpected. Even though both the chi-squared test and Fisher's Exact Test showed very low p-values (0.0000000001414 and 0.000000000003261), indicating a significant difference between the groups, the higher infection rate with the vaccine suggests we need to look into these surprising results further. Because of this increased infection risk, we end up failing to reject the null hypothesis.

Chi-Square Test p-value: 1.414

 
Fisher’s Exact Test p-value: 3.261

Odds Ratio: 4.27 (95% CI: [2.22, 8.88])


# Overall Results and Conclusions

The Covid921 vaccine has been shown to significantly reduce Covid921 cases among males, females, and the LGBTQ community, with extremely low P-values consistently rejecting the null hypothesis and demonstrating a strong protective effect compared to the placebo. However, for drug users, the data unexpectedly shows a higher risk of infection, despite also having a low P-value. This indicates a need for further investigation to uncover any potential data issues or unique factors affecting drug users. While the vaccine is recommended for the general population, additional caution and analysis are necessary for drug users.




