---
title: "Covid91 vaccine study Final2024"
author: "Mariahlynn"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: " summer 2024" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$: The Covid921 vaccine is not effective in males
$H_A$: The Covid921 vaccine is not effective in females 

### Methods

Treatment is a categorical variable  
MALES is also a categorical variable  
We use the cat~cat helper file  

### Descriptive Results 

### Graphical Results 
```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="Percent")
```

This bar charts shows that MALES who received the vaccine were less likely to get Covid921 in comparison to those who received a placebo. Therefore about 50% of MALES who received either the vaccine or placebo did not contract the virus. 


### Numerical Results 
```{r}
table1 <- xtabs(~treatment + infected, data=males)
rowPerc(table1)
colPerc(table1)
```

This shows that 74% of the males who took the placebo got covid19, as were 26% males who recieved the vaccine got the virus. 

### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi-squared test and Fisher's Exact Test both show a P-value of 0.00000000000000022, which is less than 0.05. 
Therefore, we reject the null hypothesis. A statistically significant effect of the vaccine.  We see that MALES are 2.8 times more likely to contract Covid921 with the placebo compared to the vaccine.

## LGBTQ

$H_0$: The Covid921 vaccine is not effective in LGBTQ Community.

$H_A$: The Covid921 vaccine is effective in LGBTQ Community.


### Methods 

Treatment is a categorical variable. LGBTQ is a categorical variable,so we use is cat~cat helper file    

### Descriptive Results 

### Graphical Results

```{r}
LGBTQ = filter(FinalData, LGBTQ=="gay")

barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")
```

The bar charts illustrate that LGBTQ individuals who were vaccinated had a lower likelihood of contracting Covid-19 compared to those who received a placebo. Nevertheless, approximately 45% of the LGBTQ individuals who received the placebo remained uninfected, while around 50% of those who received the vaccine were uninfected.   


### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

The numerical data reveal that roughly 80% of LGBTQ participants who received the placebo contracted Covid-19, compared to only 20% of those who received the vaccine. This indicates that the vaccine lowers the relative risk of infection by about 55.2%.

## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions